---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library(plyr)
library(timeDate)
activity<-read.csv(unz('activity.zip','activity.csv'))
activity_clean<-activity[complete.cases(activity),]
temp<-write.csv(activity_clean,'clean_activity.csv')
activity_clean<-read.csv('clean_activity.csv',row.names=1)

```

## What is mean total number of steps taken per day?

The histogram of activity is below
```{r}
z<-tapply(activity_clean$steps,activity_clean$date,sum)
hist(z,20,freq=TRUE)
```

The Mean of the number of steps is below
```{r}
mean(z)
```

Median is:
```{r}
median(z)
```
## What is the average daily activity pattern?

The activity pattern looks as follows:
```{r}
z<-tapply(activity_clean$steps,activity_clean$interval,sum)
plot(z,type='l')
```

The most active 5 minute interval on average is:
```{r}
names(which.max(z))
```


## Imputing missing values

We compute the missing values based on the median data. The number of missing values are:
```{r}
sum(is.na(activity$steps))
```

We store the median from the clean data source for the 5 minute interval and look up as necessary.
```{r}
z<-tapply(activity_clean$steps,activity_clean$interval,median)
activity_patch<-ddply(activity,.(date,interval),mutate,new_steps=if(is.na(steps)) as.numeric(z[toString(interval)]) else steps)
z<-tapply(activity_patch$new_steps,activity_patch$date,sum)
hist(z,20,freq=TRUE)
```

Mean is
```{r}
mean(z)
```

Median is
```{r}
median(z)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
activity_alldays<-ddply(activity_patch,.(date,interval),mutate,weekdayinfo= isWeekday(as.Date(date)))
activity_alldays<-ddply(activity_alldays,.(date,interval),mutate,new_interval= interval %/% 100 * 100 + ((interval %% 100)/60*100))
activity_weekdays<-ddply(activity_alldays,.(new_interval,weekdayinfo),summarize,total=sum(new_steps))
```

Plot is below. There are differences in activity patterns. As you can wee there are more steps in weekend than on weekdays.

```{r,fig.width=8,fig.height=10}
par(mfrow=c(2,1))
plot(activity_weekdays[activity_weekdays$weekdayinfo==FALSE,][,c("new_interval","total")], main="Weekend")
plot(activity_weekdays[activity_weekdays$weekdayinfo==TRUE,][,c("new_interval","total")], main="Weekday")
```